Python to Assembly Compilation

### 2011-12-12 paper/paper.tex
talk about list addition problems and our change to program syntax by forcing an assignment

\subsection{Syntax Tree Modifications}
We added two new stages to our reference compiler.  The first stage is a modified flattening stage, which produces a flattened AST to aid in the reference determination.  The next stage is the actual reference determination and AST transformation, where the compiler determines what variables will be objects and when assignments happen and adds appropriate calls to increment and decrement the reference counter of the object as needed.
\par
Our initial assumption was that we could easily add the increment and decrement reference counter operations in the reference compilers flatten stage, however because this stage is after type explication and there are variables created which do not correspond to a type in the language the pre-flatten stage was necessary.  This stage converts all expressions in to assignments to temporary variables.  At this stage we know every expression is operating on types.
\par
The next new stage is the reference counting AST transformation.  In this stage we take the flattened tree and look for assignments which are setting list elements, dictionary values, class attributes, object attributes or variables.  The compiler determines what variables are created in the current scope and injects code at the beginning of that scope to set the value initially.  We decided on this approach to simplify determining when an object is first assigned and this allows us to simply decrement the reference counter to the object before an assign, regardless of the variables state.
\subsection{Runtime Modifications}
Our runtime consisted of C functions which took care of allocating and initializing memory for a new object of a given type.  The obvious modifications included setting the reference counter initially to zero and creating a number of free functions to free each type of object.
\par
In the free\emph{\_type} functions it was necessary to add further calls to decrement the reference counter of any objects referenced by the type.  For a list list we iterate the values contained in the list, decrementing their reference (which can result in freeing that object).  For a dictionary we decrement the reference counter on keys as well as values contained in the dictionary, since both can be objects.
\subsection{Interesting Problems}

\subsection{Assembly Output}
We modified the x86 assembly to include calls to set up our memory tracking system and print statistics at the end of the program.
\section{Results}
\subsection{Execution}
\subsection{Memory}
\section{Future work}
\section{Related Work}
\appendix
\section{Appendix Title}

>>>>

\subsection{Syntax Tree Modifications}
We added two new stages to our reference compiler.  The first stage is a modified flattening stage, which produces a flattened AST to aid in the reference determination.  The next stage is the actual reference determination and AST transformation, where the compiler determines what variables will be objects and when assignments happen and adds appropriate calls to increment and decrement the reference counter of the object as needed.
Our initial assumption was that we could easily add the increment and decrement reference counter operations in the reference compilers flatten stage, however because this stage is after type explication and there are variables created which do not correspond to a type in the language the pre-flatten stage was necessary.  This stage converts all expressions in to assignments to temporary variables.  At this stage we know every expression is operating on types.
The next new stage is the reference counting AST transformation.  In this stage we take the flattened tree and look for assignments which are setting list elements, dictionary values, class attributes, object attributes or variables.  The compiler determines what variables are created in the current scope and injects code at the beginning of that scope to set the value initially.  We decided on this approach to simplify determining when an object is first assigned and this allows us to simply decrement the reference counter to the object before an assign, regardless of the variables state.
\subsection{Runtime Modifications}
Our runtime consisted of C functions which took care of allocating and initializing memory for a new object of a given type.  The obvious modifications included setting the reference counter initially to zero and creating a number of free functions to free each type of object.
In the free\emph{\_type} functions it was necessary to add further calls to decrement the reference counter of any objects referenced by the type.  For a list list we iterate the values contained in the list, decrementing their reference (which can result in freeing that object).  For a dictionary we decrement the reference counter on keys as well as values contained in the dictionary, since both can be objects.
\subsection{Interesting Problems}
We encountered a number of interesting questions during this implementation and summarize them below.
\subsubsection{Addition of Lists}
In the current runtime an entirely new list is created when two lists are added together.  At the assignment level our compiler can determine (due to the flattened structure) that the left-hand-side and the right-hand-side of the addition need their reference counters manipulated, however the assignment from an addition to a list creates an entirely new list which should properly increment the reference counter to all objects in both lists.
\subsubsection{Variable Assignment}
We wanted a simple way to determine when to decrement an objects reference counter.  Our first implementation was to add a call to decrement the counter before every assignment.  However, because a variable can be created at any point in program execution we couldn't just blindly decrement the counter of an object since the variable will be uninitialized, and could be pointing at memory that makes it look like an object or accidentally references an object.
The solution we took makes a slight modification to program behaviour, but we feel it is an acceptable trade-off.  In every scope, we add an immediate assignment as the first instructions.  This assignment is to a primitive, so it costs us no memory and then a decrement on the reference counter amounts to a function call with a no-op.  This function call overhead could potentially be avoided by adding code to check if the variable points to an object or a primitive and is a future enhancement for performance.
\subsection{Assembly Output}
We modified the x86 assembly to include calls to set up our memory tracking system and print statistics at the end of the program.
\section{Results}
\subsection{Execution}
\subsection{Memory}
\section{Future work}
\section{Related Work}
\appendix
\section{Appendix Title}
